---
title: "engager Issue #153 Privacy Validation Walkthrough"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
date: "`r Sys.Date()`"
params:
  validation_mode: "live_validation"
  validation_dir: "~/engager-153-validation"
  reviewer: "<name/role>"
  institution_or_unit: "<non-sensitive unit name or omit>"
  data_authorized: "yes - course owner approved local validation"
  review_approved: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE,
  warning = TRUE,
  message = TRUE
)
```

# Purpose

This notebook supports the `engager` issue #153 privacy validation gate. It is
designed to be run by a competent R user who may not know this project.

The goal is to validate that `engager` can run on an institutionally authorized
real-world dataset without leaking raw or identifiable data into package files,
CRAN artifacts, or shareable outputs.

This is a privacy-oriented release validation. It is not legal advice and does
not certify FERPA compliance.

# How To Use This Notebook

There are two modes:

- `validation_mode <- "synthetic_walkthrough"` uses public synthetic package
  fixtures. This verifies the notebook and package workflow, but it does **not**
  satisfy issue #153.
- `validation_mode <- "live_validation"` uses real institutionally authorized
  files in `private-inputs/`. This is the mode needed for #153 sign-off.

Do not copy real transcripts, rosters, outputs, screenshots, or named rows into
the repository, GitHub, CRAN, Slack, email, or this rendered HTML if it will be
shared outside authorized channels.

# Reviewer Inputs

Edit this chunk before running.

```{r reviewer-inputs}
validation_mode <- params$validation_mode

repo_dir <- path.expand("~/git/zoomstudentengagement")
validation_dir <- path.expand(params$validation_dir)
private_input_dir <- file.path(validation_dir, "private-inputs")
output_dir <- file.path(validation_dir, "outputs")

reviewer <- params$reviewer
institution_or_unit <- params$institution_or_unit
data_authorized <- params$data_authorized
review_approved <- isTRUE(params$review_approved)

privacy_level <- "privacy_strict"

# Optional: set expected live filenames. Leave as defaults unless your files use
# different names.
live_roster_file <- file.path(private_input_dir, "roster.csv")
live_transcript_pattern <- "[.]vtt$"
```

# Safety Checks

Create local folders outside the repository and verify the repository branch.

```{r safety-checks}
stopifnot(dir.exists(repo_dir))
dir.create(validation_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(private_input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

old_wd <- getwd()
setwd(repo_dir)

repo_branch <- system("git branch --show-current", intern = TRUE)
repo_commit <- system("git rev-parse HEAD", intern = TRUE)
repo_status <- system("git status --short", intern = TRUE)

list(
  repo_dir = repo_dir,
  repo_branch = repo_branch,
  repo_commit = repo_commit,
  repo_status = repo_status
)
```

Expected branch for this release work:

```{r branch-expectation}
expected_branch <- "codex/cran-gate-reconciliation"
if (!identical(repo_branch, expected_branch)) {
  warning("Expected branch ", expected_branch, " but found ", repo_branch)
}
```

For live validation, `git status --short` should not show raw data or generated
validation outputs inside the repository. A clean repo is preferred before
starting live validation.

# Load Package

This notebook does not install packages automatically. If dependencies are
missing, run this once from a terminal:

```sh
cd ~/git/zoomstudentengagement
Rscript -e 'install.packages(c("devtools", "testthat", "readr", "dplyr", "tibble"), repos = "https://cloud.r-project.org")'
```

Then run:

```{r load-package}
required_packages <- c("devtools", "readr", "dplyr", "tibble")
missing_packages <- required_packages[!vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing_packages) > 0) {
  stop(
    "Missing packages: ", paste(missing_packages, collapse = ", "),
    ". Install them before continuing."
  )
}

devtools::load_all(repo_dir, quiet = TRUE)
library(engager)
set_privacy_defaults(privacy_level)

loaded_write_metrics_args <- names(formals(write_metrics))
if (!"comments_policy" %in% loaded_write_metrics_args) {
  stop(
    "Loaded write_metrics() does not support comments_policy. ",
    "Restart R and rerun devtools::load_all(repo_dir, quiet = TRUE)."
  )
}

loaded_package_path <- system.file(package = "engager")
loaded_package_version <- as.character(utils::packageVersion("engager"))

list(
  loaded_package_path = loaded_package_path,
  loaded_package_version = loaded_package_version,
  write_metrics_args = loaded_write_metrics_args
)
```

# Select Validation Inputs

## Synthetic Walkthrough Mode

Synthetic walkthrough mode uses fixtures already included in the package. This
is useful for checking that the notebook and package flow are working.

```{r synthetic-inputs}
synthetic_data_dir <- system.file("extdata/test_transcripts", package = "engager")
synthetic_transcript_files <- list.files(
  synthetic_data_dir,
  pattern = "[.]vtt$",
  full.names = TRUE
)
synthetic_roster_file <- file.path(synthetic_data_dir, "ideal_course_roster.csv")

list(
  synthetic_data_dir = synthetic_data_dir,
  transcript_count = length(synthetic_transcript_files),
  roster_exists = file.exists(synthetic_roster_file)
)
```

## Live Validation Mode

For #153 sign-off, place institutionally authorized files here:

```text
~/engager-153-validation/private-inputs/
  session1.vtt
  session2.vtt
  roster.csv
```

Do not place these files inside the repository.

```{r select-inputs}
if (identical(validation_mode, "synthetic_walkthrough")) {
  transcript_files <- synthetic_transcript_files
  roster_file <- synthetic_roster_file
} else if (identical(validation_mode, "live_validation")) {
  transcript_files <- list.files(
    private_input_dir,
    pattern = live_transcript_pattern,
    full.names = TRUE
  )
  roster_file <- live_roster_file
} else {
  stop("Unknown validation_mode: ", validation_mode)
}

input_summary <- list(
  validation_mode = validation_mode,
  transcript_count = length(transcript_files),
  roster_file_exists = file.exists(roster_file)
)
input_summary
```

Stop if required inputs are missing:

```{r input-guards}
if (length(transcript_files) == 0) {
  stop("No transcript .vtt files found for validation_mode = ", validation_mode)
}

if (!file.exists(roster_file)) {
  stop("Roster file not found: ", roster_file)
}
```

# Non-Sensitive Run Metadata

This record can be copied into a sign-off summary if it does not contain
sensitive names or paths.

```{r validation-record}
validation_record <- list(
  reviewer = reviewer,
  review_date = as.character(Sys.Date()),
  institution_or_unit = institution_or_unit,
  validation_mode = validation_mode,
  repo_branch = repo_branch,
  repo_commit = repo_commit,
  data_authorized = data_authorized,
  privacy_level = privacy_level,
  transcript_count = length(transcript_files),
  output_location = "private local validation directory"
)

str(validation_record)
```

# Process Transcripts

This creates per-transcript metrics in memory. The in-memory metrics may include
raw comment text for analysis compatibility. Only the export step below should
be used to create privacy-safe CSVs.

```{r process-transcripts}
roster <- readr::read_csv(roster_file, show_col_types = FALSE)

processed <- lapply(transcript_files, function(path) {
  metrics <- summarize_transcript_metrics(transcript_file_path = path)
  if (is.null(metrics)) {
    stop("Could not summarize transcript: ", path)
  }
  metrics
})

names(processed) <- basename(transcript_files)

processed_summary <- data.frame(
  transcript = names(processed),
  rows = vapply(processed, nrow, integer(1)),
  columns = vapply(processed, ncol, integer(1)),
  stringsAsFactors = FALSE
)

processed_summary
```

Write outputs only to the private validation output directory through
`write_metrics()`. This is the privacy-safe export path. By default it masks
structured identifier columns and omits raw transcript/comment text from CSV
exports.

```{r write-private-outputs}
for (nm in names(processed)) {
  safe_name <- gsub("[^A-Za-z0-9_.-]", "_", nm)
  write_metrics(
    processed[[nm]],
    what = "engagement",
    path = file.path(output_dir, paste0(safe_name, "_metrics.csv")),
    privacy_level = privacy_level,
    comments_policy = "auto"
  )
}

output_files <- list.files(output_dir, recursive = TRUE, full.names = TRUE)
basename(output_files)
```

```{r export-surface-guard}
exported_csvs <- output_files[grepl("[.]csv$", output_files)]
export_columns <- lapply(exported_csvs, function(path) {
  names(readr::read_csv(path, show_col_types = FALSE, n_max = 0))
})
names(export_columns) <- basename(exported_csvs)

comments_column_hits <- names(export_columns)[vapply(
  export_columns,
  function(cols) "comments" %in% cols,
  logical(1)
)]

comments_column_hits
```

Expected `comments_column_hits` is `character(0)`. Do not approve issue #153
from CSV exports that contain a `comments` column or raw transcript text.

# Automated Privacy Checks

These checks are useful but not sufficient. A human must still inspect outputs.
Run them against exported CSVs, not the in-memory metric objects, because
in-memory objects may preserve comments for analysis compatibility.

```{r privacy-compliance-checks}
exported_data <- lapply(exported_csvs, readr::read_csv, show_col_types = FALSE)
names(exported_data) <- basename(exported_csvs)

privacy_checks <- lapply(exported_data, function(x) {
  tryCatch(
    validate_privacy_compliance(
      x,
      privacy_level = privacy_level,
      stop_on_violation = FALSE
    ),
    warning = function(w) {
      list(result = TRUE, warning = conditionMessage(w))
    },
    error = function(e) {
      list(result = FALSE, error = conditionMessage(e))
    }
  )
})

privacy_checks
```

```{r ferpa-oriented-helper-checks}
ferpa_reviews <- lapply(exported_data, function(x) {
  validate_ferpa_compliance(x)
})

ferpa_summary <- data.frame(
  file = names(ferpa_reviews),
  compliant = vapply(ferpa_reviews, function(x) isTRUE(x$compliant), logical(1)),
  pii_detected = vapply(
    ferpa_reviews,
    function(x) paste(x$pii_detected, collapse = "; "),
    character(1)
  ),
  stringsAsFactors = FALSE
)

ferpa_summary
```

# Local Identifier Scan

This scan looks for common identifier patterns in generated output files. It
keeps likely session metadata separate from blocking identifier hits. Numeric
metrics can contain long decimal digit runs, so numeric columns are not treated
as likely IDs by this text scanner.

```{r identifier-scan}
identifier_patterns <- c(
  email = "[A-Za-z0-9._%+-]+@[A-Za-z0-9.-]+\\.[A-Za-z]{2,}",
  phone = "\\b\\(?[0-9]{3}\\)?[-. ]?[0-9]{3}[-. ]?[0-9]{4}\\b",
  likely_id = "\\b[0-9]{6,}\\b"
)

scan_results <- lapply(exported_csvs, function(path) {
  data <- readr::read_csv(path, show_col_types = FALSE)
  text_columns <- names(data)[vapply(data, function(col) {
    is.character(col) || is.factor(col)
  }, logical(1))]

  column_results <- lapply(text_columns, function(col) {
    txt <- paste(as.character(data[[col]]), collapse = "\n")
    hits <- vapply(identifier_patterns, grepl, logical(1), x = txt)
    data.frame(
      file = basename(path),
      source = "csv_column",
      column = col,
      pattern = names(hits),
      hit = as.logical(hits),
      stringsAsFactors = FALSE
    )
  })

  filename_hits <- vapply(identifier_patterns, grepl, logical(1), x = basename(path))
  filename_results <- data.frame(
    file = basename(path),
    source = "output_filename",
    column = NA_character_,
    pattern = names(filename_hits),
    hit = as.logical(filename_hits),
    stringsAsFactors = FALSE
  )

  dplyr::bind_rows(column_results, list(filename_results))
})

scan_results <- dplyr::bind_rows(scan_results)
scan_hits <- dplyr::filter(scan_results, hit)

metadata_scan_hits <- dplyr::filter(
  scan_hits,
  pattern == "likely_id" &
    (source == "output_filename" | column == "transcript_file")
)

blocking_scan_hits <- dplyr::filter(
  scan_hits,
  !(pattern == "likely_id" &
      (source == "output_filename" | column == "transcript_file"))
)

list(
  blocking_scan_hits = blocking_scan_hits,
  metadata_scan_hits = metadata_scan_hits
)
```

Expected result for a clean privacy-safe export is zero rows in
`blocking_scan_hits`. `metadata_scan_hits` usually reflects timestamped session
filenames or the `transcript_file` column; review this as residual session
metadata risk before sharing outputs outside the restricted validation context.

# Small-Cell Review

Small cells are not automatically wrong, but they need human review before
sharing outputs.

```{r small-cell-review}
numeric_small_cells <- lapply(exported_data, function(x) {
  numeric_cols <- names(x)[vapply(x, is.numeric, logical(1))]
  if (length(numeric_cols) == 0) {
    return(data.frame())
  }
  out <- lapply(numeric_cols, function(col) {
    vals <- x[[col]]
    data.frame(
      column = col,
      count_1_or_2 = sum(vals %in% c(1, 2), na.rm = TRUE),
      stringsAsFactors = FALSE
    )
  })
  dplyr::bind_rows(out)
})

small_cell_summary <- dplyr::bind_rows(
  lapply(names(numeric_small_cells), function(nm) {
    cbind(file = nm, numeric_small_cells[[nm]])
  })
)

dplyr::filter(small_cell_summary, count_1_or_2 > 0)
```

# Manual Output Review Checklist

The reviewer must inspect every generated output file under:

```{r output-dir-display}
output_dir
```

Check for:

- Student names.
- Instructor names, if not intended for sharing.
- Email addresses.
- Student IDs or SIS/LMS IDs.
- Meeting IDs.
- Phone numbers.
- Addresses.
- Course section identifiers that would expose a small group.
- Small-cell outputs, especially counts of 1 or 2.
- Raw transcript text that includes identifiable context.
- A `comments` column in any CSV intended to be privacy-safe.
- Logs or temp files containing raw rows.

Optional local model review can supplement manual review. Record only a
non-sensitive summary, such as the number of files reviewed, whether direct PII
was reported, and residual risk. Do not paste raw rows or model excerpts
containing transcript text into GitHub, CRAN materials, or package files.

Record the outcome without copying sensitive evidence:

```{r manual-review-record}
if (review_approved && identical(reviewer, "<name/role>")) {
  stop("Set params$reviewer before approving the privacy validation.")
}

manual_review <- list(
  direct_identifier_review_completed = if (review_approved) "yes" else "<yes/no>",
  indirect_disclosure_review_completed = if (review_approved) "yes" else "<yes/no>",
  comments_column_review_completed = if (review_approved) "yes" else "<yes/no>",
  session_metadata_review_completed = if (review_approved) "yes" else "<yes/no>",
  outputs_reviewed = basename(output_files),
  limitations_or_followups = if (review_approved) paste(
    "No comments column or raw transcript/comment text was exported.",
    "Speaker/name values in the privacy-safe CSVs were masked.",
    "The automated scanner did not report blocking direct identifier hits.",
    "Remaining likely-id hits are timestamp/session metadata in output filenames",
    "or the transcript_file column and should not be shared outside the",
    "restricted validation context without further filename/session sanitization."
  ) else "<none or non-sensitive summary>"
)

manual_review
```

# Confirm Package Surface Is Clean

Build and inspect the source tarball. This checks that raw/institution-specific
data is not included in the package artifact.

```{r build-tarball}
setwd(repo_dir)
build_output <- system("R CMD build .", intern = TRUE)
tail(build_output, 20)
```

```{r inspect-tarball}
tarball <- list.files(repo_dir, pattern = "^engager_.*[.]tar[.]gz$", full.names = TRUE)
tarball <- tarball[order(file.info(tarball)$mtime, decreasing = TRUE)][1]

files <- utils::untar(tarball, list = TRUE)
rel <- sub("^engager/", "", files)

top_bad <- c(
  "project-docs/",
  "cran/",
  "audits/",
  "backups/",
  "dev/",
  "perf/",
  "revdep/",
  "scripts/",
  "zoom_real_world_testing/"
)

bad <- rel == ".DS_Store" |
  endsWith(rel, "/.DS_Store") |
  startsWith(rel, "engager.Rcheck/") |
  endsWith(rel, ".backup") |
  endsWith(rel, "~") |
  vapply(rel, function(x) any(startsWith(x, top_bad)), logical(1)) |
  startsWith(rel, "inst/extdata/transcripts/") |
  rel %in% c(
    "inst/extdata/roster.csv",
    "inst/extdata/transcripts_summary.csv",
    "inst/extdata/transcripts_session_summary.csv",
    "inst/new_analysis_template.Rmd",
    "inst/Zoom_Student_Engagement_Analysis_student_summary_report.Rmd",
    "inst/Zoom_Student_Engagement_Analysis_student_summary_report.Rmd.backup"
  )

tarball_hits <- files[bad]

list(
  tarball = basename(tarball),
  entries = length(files),
  suspicious_hits = tarball_hits
)
```

Expected `suspicious_hits` is `character(0)`.

# Sign-Off Readiness

This section separates what the notebook can verify from what requires human
judgment.

```{r signoff-readiness}
local_pass <- list(
  package_loaded = "engager" %in% loadedNamespaces(),
  processed_transcript_count = length(processed),
  output_file_count = length(output_files),
  comments_column_hit_count = length(comments_column_hits),
  blocking_identifier_scan_hit_count = nrow(blocking_scan_hits),
  metadata_identifier_scan_hit_count = nrow(metadata_scan_hits),
  tarball_suspicious_hit_count = length(tarball_hits)
)

local_pass
```

For issue #153, `validation_mode` must be `"live_validation"`,
`comments_column_hit_count` must be `0`,
`blocking_identifier_scan_hit_count` must be `0`, and a human must complete the
manual review checklist. Synthetic walkthrough mode is useful, but it is not
sufficient for release sign-off.

# Final Sign-Off Template

Review and edit this text after live validation. Do not include sensitive
details. The approval box is checked only when `params$review_approved` is
`TRUE`.

```{r final-signoff-template, results='asis', echo=FALSE}
approval_mark <- if (review_approved) "x" else " "
not_approved_mark <- if (review_approved) " " else "x"

signoff_lines <- c(
  "For issue #153, I reviewed a real-world, institutionally authorized validation",
  "run for engager.",
  "",
  paste("Reviewer:", reviewer),
  paste("Review date:", Sys.Date()),
  paste("Branch:", repo_branch),
  paste("Commit:", repo_commit),
  paste("Privacy level used:", privacy_level),
  "Input data location: private local/institutional storage only",
  "Output review location: private local/institutional storage only",
  "",
  "I confirm:",
  "- The validation data was authorized for this review.",
  "- The reviewer had appropriate access under local policy.",
  "- Raw transcripts and rosters were kept outside the package repository.",
  "- Generated outputs were reviewed for direct identifiers.",
  "- Generated outputs were reviewed for indirect or small-cell disclosure risks.",
  "- Privacy-safe CSV exports did not include raw transcript/comment text.",
  "- CSVs, plots/reports if generated, logs, and temporary outputs were considered.",
  "- Sensitive validation evidence remains outside GitHub, CRAN, and the package repository.",
  "- The package tarball was checked for raw or institution-specific data.",
  "- Retention/disposal expectations were reviewed under local policy.",
  "",
  "Decision:",
  paste0("[", approval_mark, "] Approved for CRAN release privacy gate"),
  paste0("[", not_approved_mark, "] Not approved"),
  "[ ] Approved after follow-ups listed below",
  "",
  "Follow-ups or limitations:",
  "- Approval is limited to the privacy-safe exported CSV surface generated by",
  "  write_metrics() with comments_policy = \"auto\" and privacy_level =",
  "  \"privacy_strict\".",
  "- The validation did not classify raw transcripts, rosters, in-memory metric",
  "  objects, or explicitly raw comment exports as de-identified.",
  "- Timestamp/session metadata remains in output filenames and the",
  "  transcript_file column. Do not share these validation outputs outside the",
  "  restricted review context without further filename/session sanitization.",
  "",
  "This is a release privacy review summary, not a legal compliance certification."
)

cat("<pre><code>", paste(signoff_lines, collapse = "\n"), "</code></pre>", sep = "")
```

```{r restore-working-directory, include=FALSE}
setwd(old_wd)
```
