---
title: "Essential Functions - Core Package Functionality"
author: "zoomstudentengagement package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Essential Functions - Core Package Functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This vignette demonstrates the core functionality of the `zoomstudentengagement` package using only the essential functions. After scope reduction, the package now exports **26 essential functions** focused on:

- **Transcript Processing**: Loading and processing Zoom transcripts
- **Privacy Protection**: Ensuring FERPA compliance and data privacy
- **Analysis**: Core engagement metrics and analysis
- **Name Matching**: Safe student name matching workflows

## Essential Functions

The package now exports these 26 essential functions:

```{r essential-functions}
library(zoomstudentengagement)

# Get list of essential functions
essential_functions <- get_essential_functions()
cat("Essential functions (", length(essential_functions), "):\n")
cat(paste("-", essential_functions), sep = "\n")
```

## Core Workflow

### 1. Load and Process Transcript

```{r load-transcript}
# Load a sample transcript
transcript_file <- system.file("extdata/transcripts/GMT20240124-202901_Recording.transcript.vtt",
  package = "zoomstudentengagement"
)

# Load the transcript
transcript <- load_zoom_transcript(transcript_file)
cat("Transcript loaded with", nrow(transcript), "comments\n")
```

### 2. Process with Privacy Protection

```{r process-transcript}
# Process transcript with privacy defaults
processed <- process_zoom_transcript(
  transcript_df = transcript,
  add_dead_air = TRUE,
  consolidate_comments = TRUE
)

cat("Processed transcript has", nrow(processed), "rows\n")
```

### 3. Analyze Engagement

```{r analyze-engagement}
# Analyze transcripts for engagement metrics
analysis <- analyze_transcripts(
  transcript_files = transcript_file,
  roster_file = NULL,
  privacy_level = "FERPA"
)

cat("Analysis completed successfully\n")
```

### 4. Privacy and Compliance

```{r privacy-compliance}
# Check privacy compliance
privacy_status <- validate_privacy_compliance(processed)
cat("Privacy compliance:", privacy_status, "\n")

# FERPA compliance check
ferpa_status <- validate_ferpa_compliance(processed)
cat("FERPA compliance:", ferpa_status, "\n")
```

## Function Categories

### Transcript Processing
- `load_zoom_transcript()` - Load Zoom transcript files
- `process_zoom_transcript()` - Process and clean transcripts
- `consolidate_transcript()` - Consolidate consecutive comments
- `add_dead_air_rows()` - Add gaps between comments

### Analysis
- `analyze_transcripts()` - Main analysis function
- `summarize_transcript_metrics()` - Calculate engagement metrics
- `analyze_multi_session_attendance()` - Multi-session analysis
- `generate_attendance_report()` - Generate attendance reports

### Privacy & Compliance
- `ensure_privacy()` - Apply privacy protection
- `set_privacy_defaults()` - Configure privacy settings
- `privacy_audit()` - Audit privacy compliance
- `validate_privacy_compliance()` - Validate privacy settings
- `validate_ferpa_compliance()` - FERPA compliance check

### Name Matching
- `load_roster()` - Load student roster
- `safe_name_matching_workflow()` - Safe name matching
- `detect_unmatched_names()` - Find unmatched names
- `detect_duplicate_transcripts()` - Find duplicate files

### Utilities
- `plot_users()` - Plot user engagement
- `write_metrics()` - Export metrics
- `validate_schema()` - Validate data structure

## Benefits of Scope Reduction

1. **Focused Functionality**: Only essential functions are exported
2. **Easier Maintenance**: Smaller codebase to maintain
3. **Better User Experience**: Clear, focused API
4. **CRAN Ready**: Minimal function surface for submission
5. **Privacy First**: All functions prioritize data protection

## Next Steps

With the scope reduction complete, the package is now ready for:

1. **CRAN Submission**: Minimal, focused function set
2. **User Testing**: Clear, essential functionality
3. **Documentation Updates**: Focus on core workflows
4. **UX Simplification**: Streamlined user experience

The package now provides a clean, focused API for analyzing student engagement from Zoom transcripts while maintaining strong privacy protection and FERPA compliance.
