---
title: "Getting Started with zoomstudentengagement"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Getting Started with zoomstudentengagement

The `zoomstudentengagement` package makes it easy to analyze student engagement from Zoom transcripts. This guide will get you started with the basics in just a few minutes.

## Quick Start

The easiest way to analyze Zoom transcripts is with the basic workflow:

```{r quick-start}
# Load the package
library(zoomstudentengagement)

# Run basic analysis (recommended for new users)
results <- basic_transcript_analysis("your_transcript.vtt", "output/")
```

That's it! The analysis will:
1. Load your transcript file
2. Process and clean the data
3. Calculate engagement metrics
4. Create visualizations
5. Export results to files

## Step-by-Step Workflow

If you prefer more control, you can run each step individually:

### 1. Load Transcript
```{r load-transcript}
transcript <- load_zoom_transcript("your_transcript.vtt")
```

### 2. Process Data
```{r process-data}
processed <- process_zoom_transcript(transcript)
```

### 3. Analyze Engagement
```{r analyze-engagement}
analysis <- analyze_transcripts(processed)
```

### 4. Create Visualizations
```{r create-visualizations}
plots <- plot_users(analysis)
```

### 5. Export Results
```{r export-results}
write_metrics(analysis, "output/")
```

## Supported File Formats

The package supports these Zoom transcript formats:
- **VTT files** (`.vtt`) - Video transcript files
- **TXT files** (`.txt`) - Text transcript files  
- **CSV files** (`.csv`) - Comma-separated transcript data

## Privacy Protection

Privacy protection is enabled by default. The package automatically:
- Protects student names and personal information
- Applies FERPA compliance features
- Provides data anonymization options

```{r privacy-example}
# Privacy protection is automatic, but you can configure it
set_privacy_defaults()  # Use default privacy settings
ensure_privacy(data)    # Apply privacy protection to data
```

## Getting Help

The package includes several help functions to guide you:

```{r help-examples}
# Show getting started guide
show_getting_started()

# See available functions at your level
show_available_functions()

# Get help for a specific function
show_function_help("load_zoom_transcript")

# Find the right function for your task
find_function_for_task("create visualizations")
```

## Progressive Disclosure

The package uses progressive disclosure to keep the interface simple:

- **Basic level** (default): 5 essential functions
- **Intermediate level**: 15 common functions  
- **Advanced level**: 35+ specialized functions
- **Expert level**: All 79 functions

```{r progressive-disclosure}
# Start with basic level (default)
show_available_functions("basic")

# See more functions
set_ux_level("intermediate")
show_available_functions()

# Access all functions
set_ux_level("expert")
```

## Batch Processing

For multiple transcript files:

```{r batch-processing}
# Process multiple files at once
files <- c("session1.vtt", "session2.vtt", "session3.vtt")
results <- batch_basic_analysis(files, "batch_output/")
```

## Common Workflows

### Quick Analysis
```{r quick-analysis}
# Fastest way to get results
results <- quick_analysis("transcript.vtt")
```

### Privacy-Focused Analysis
```{r privacy-analysis}
# Maximum privacy protection
set_privacy_defaults()
results <- basic_transcript_analysis("transcript.vtt", "output/", "high")
```

### Custom Analysis
```{r custom-analysis}
# Step-by-step with custom options
transcript <- load_zoom_transcript("transcript.vtt")
processed <- process_zoom_transcript(transcript)
analysis <- analyze_transcripts(processed)
plots <- plot_users(analysis)
write_metrics(analysis, "custom_output/")
```

## Troubleshooting

If you encounter issues:

```{r troubleshooting}
# Show troubleshooting guide
show_troubleshooting()

# Get specific help
show_function_help("function_name")

# Find the right function
find_function_for_task("what you want to do")
```

## Next Steps

1. **Try the basic workflow** with your transcript files
2. **Explore more functions** with `set_ux_level("intermediate")`
3. **Check the privacy features** with `show_privacy_guidance()`
4. **Learn advanced techniques** in the advanced vignettes

## Resources

- **Package documentation**: `help(package = "zoomstudentengagement")`
- **Function help**: `show_function_help("function_name")`
- **Getting started**: `show_getting_started()`
- **Privacy guide**: `show_privacy_guidance()`
- **Troubleshooting**: `show_troubleshooting()`

Happy analyzing! 🎯
