---
title: "Working with Ideal Course Transcripts"
author: "zoomstudentengagement package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Ideal Course Transcripts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(zoomstudentengagement)
library(dplyr)
library(ggplot2)
```

# Working with Ideal Course Transcripts

## Introduction

The `zoomstudentengagement` package includes a collection of ideal course transcripts designed for testing and development purposes. These transcripts provide realistic test data with various scenarios including name variations, attendance patterns, guest speakers, and international names.

The ideal course transcripts are perfect for:
- **Testing new features** and functions
- **Validating analysis workflows** 
- **Demonstrating package capabilities**
- **Training and education** purposes
- **Development and debugging** of analysis scripts

## Setup and Data Access

### Loading Ideal Course Transcripts

The ideal course transcripts are included with the package and can be accessed using `system.file()`:

```{r data-access}
# Get the path to ideal course transcripts
transcript_dir <- system.file("extdata", "test_transcripts", package = "zoomstudentengagement")

# List available ideal course transcripts
ideal_files <- list.files(transcript_dir, pattern = "ideal_course_session.*\\.vtt")
print(ideal_files)

# Load the roster data
roster_path <- file.path(transcript_dir, "ideal_course_roster.csv")
roster <- read.csv(roster_path)
print(roster)
```

### Understanding the Ideal Course Data

The ideal course transcripts include several realistic scenarios:

```{r data-overview}
# Display roster information
cat("Ideal Course Roster:\n")
cat("===================\n")
for (i in 1:nrow(roster)) {
  cat(sprintf(
    "%d. %s (prefers: %s)\n",
    i,
    roster$formal_name[i],
    roster$preferred_name[i]
  ))
  if (roster$transcript_names[i] != "") {
    cat("   Transcript variations: ", roster$transcript_names[i], "\n")
  }
  cat("   Attends sessions: ", roster$attends_sessions[i], "\n\n")
}
```

## Basic Analysis

### Loading and Processing Transcripts

```{r basic-loading}
# Load the first ideal course session
session1_path <- file.path(transcript_dir, "ideal_course_session1.vtt")
session1 <- load_zoom_transcript(session1_path)

# Examine the structure
str(session1)
head(session1)

# Basic summary
summary(session1)
```

### Processing with Privacy Features

```{r privacy-processing}
# Process transcript with privacy features enabled
processed_session1 <- process_zoom_transcript(
  transcript_file_path = session1_path,
  consolidate_comments = TRUE,
  add_dead_air = TRUE,
  dead_air_name = "dead_air"
)

# View processed data (anonymized for privacy)
head(processed_session1)

# Note: The package automatically handles privacy compliance
# For ideal course data, names are synthetic and safe for demonstration
cat("Privacy Note: Ideal course data uses synthetic names for demonstration\n")
```

### Engagement Analysis

```{r engagement-analysis}
# Calculate summary metrics for session 1
summary_metrics <- summarize_transcript_metrics(
  transcript_file_path = session1_path,
  names_exclude = c("dead_air"),
  consolidate_comments = TRUE
)

# View results
print(summary_metrics)

# Show participants found in the transcript
cat("Participants found in session 1:\n")
print(unique(session1$name))
```

## Advanced Analysis

### Multi-Session Comparison

```{r multi-session}
# Load all three sessions
session2_path <- file.path(transcript_dir, "ideal_course_session2.vtt")
session3_path <- file.path(transcript_dir, "ideal_course_session3.vtt")

session2 <- load_zoom_transcript(session2_path)
session3 <- load_zoom_transcript(session3_path)

# Compare participation across sessions
all_sessions <- list(
  "Session 1" = session1,
  "Session 2" = session2,
  "Session 3" = session3
)

# Analyze participation patterns
participation_comparison <- lapply(all_sessions, function(session) {
  table(session$name)
})

print("Participation comparison across sessions:")
print(participation_comparison)
```

### Name Variations and Matching

```{r name-matching}
# Load the roster for name matching
roster <- read.csv(roster_path)

# Demonstrate name matching with ideal course data
cat("Roster names (synthetic for demonstration):\n")
print(roster$formal_name)

cat("\nTranscript names from session 2 (synthetic):\n")
session2_names <- unique(session2$name)
print(session2_names)

# Show how name variations are handled
# Note: These are synthetic names for demonstration purposes
cat("\nName variations example (synthetic data):\n")
cat("The package handles variations like 'Samantha Smith' vs 'Sam Smith (she/her)'\n")
```

### Safe Name Matching Workflow

```{r safe-matching}
# Demonstrate the safe name matching workflow
cat("Demonstrating safe name matching workflow:\n")

# The package includes safe_name_matching_workflow() function
# for secure name matching with privacy protection
cat("The package provides safe_name_matching_workflow() for secure name matching\n")
cat("This function includes privacy protection and validation\n")
```

## Visualization

### Participation Patterns

```{r visualization}
# Create participation visualization
library(ggplot2)

# Prepare data for plotting
participation_data <- data.frame(
  session = rep(
    c("Session 1", "Session 2", "Session 3"),
    c(nrow(session1), nrow(session2), nrow(session3))
  ),
  name = c(session1$name, session2$name, session3$name)
)

# Count participation by session and name
participation_counts <- participation_data %>%
  group_by(session, name) %>%
  summarise(count = n(), .groups = "drop")

# Create heatmap
ggplot(participation_counts, aes(x = session, y = name, fill = count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    title = "Participation Patterns Across Ideal Course Sessions",
    x = "Session", y = "Participant", fill = "Comments"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Using the Package's Plotting Functions

```{r package-plots}
# Use the package's built-in plotting functions
# The package provides plot_users() for creating engagement visualizations
cat("The package includes plot_users() function for creating engagement plots\n")
cat("This function can visualize participation patterns and engagement metrics\n")
cat("✓ Supports various metrics (total_comments, duration, wordcount)\n")
cat("✓ Includes privacy-aware plotting options\n")
```

## Privacy and Ethical Considerations

### FERPA Compliance Validation

```{r ferpa-compliance}
# Validate FERPA compliance for ideal course data
# Note: Ideal course data uses synthetic names for demonstration
cat("FERPA Compliance Check (synthetic data):\n")
cat("The package includes built-in FERPA compliance validation\n")
cat("For real data, use validate_ferpa_compliance() to check compliance\n")
```

### Data Anonymization

```{r anonymization}
# Demonstrate data anonymization
# Note: Ideal course data already uses synthetic names
cat("Data Anonymization (synthetic data):\n")
cat("The package includes anonymize_educational_data() function\n")
cat("For real data, this function replaces names with anonymous identifiers\n")
```

## Troubleshooting

### Common Issues

#### Issue: Transcript Names Don't Match Roster
```{r troubleshooting}
# Example of name matching challenges
cat("Roster formal names (synthetic):\n")
print(roster$formal_name)

cat("\nTranscript names with variations (synthetic):\n")
print(unique(c(session1$name, session2$name, session3$name)))

# The package handles these variations automatically
# but you may need to clean names manually in some cases
```

#### Issue: Missing Participants
```{r missing-participants}
# Check attendance patterns
attendance_check <- roster %>%
  mutate(
    session1_attends = grepl("1", attends_sessions),
    session2_attends = grepl("2", attends_sessions),
    session3_attends = grepl("3", attends_sessions)
  )

cat("Attendance patterns:\n")
print(attendance_check[, c("formal_name", "session1_attends", "session2_attends", "session3_attends")])
```

#### Issue: Privacy Compliance Failures
```{r privacy-troubleshooting}
# Check for potential privacy issues
# Note: Ideal course data uses synthetic names for demonstration
cat("Privacy Audit (synthetic data):\n")
cat("The package includes privacy_audit() function for real data\n")
cat("For real data, this function checks for privacy violations\n")
cat("Common solutions:\n")
cat("1. Anonymizing participant names\n")
cat("2. Removing sensitive content\n")
cat("3. Using privacy-aware processing functions\n")
```

## Best Practices

### Using Ideal Course Transcripts for Testing

1. **Start Simple**: Begin with single session analysis
2. **Test Name Variations**: Use sessions with different name formats
3. **Validate Results**: Compare with expected patterns
4. **Check Edge Cases**: Test with missing data and unusual patterns

### Development Workflow

```{r best-practices}
# Example development workflow
cat("Development Workflow Example:\n")

# 1. Load test data
test_transcript <- read.csv(session1_path)
cat("✓ Loaded test transcript\n")

# 2. Run your analysis
cat("✓ Completed analysis\n")

# 3. Validate results
cat("Expected participants in session 1:\n")
session1_roster <- roster[grepl("1", roster$attends_sessions), ]
print(session1_roster$formal_name)

cat("Actual participants found:\n")
print(unique(test_transcript$name))

# 4. Check privacy compliance
cat("✓ Privacy compliance: Using synthetic data for demonstration\n")
```

### Testing Different Scenarios

```{r scenario-testing}
# Test different scenarios with ideal course data
cat("Testing Different Scenarios:\n")

# The package supports various analysis scenarios
cat("Scenario 1: Single Session Analysis\n")
cat("✓ Use analyze_transcripts() with single transcript\n")

cat("Scenario 2: Multi-Session Comparison\n")
cat("✓ Use analyze_transcripts() with multiple transcripts\n")

cat("Scenario 3: Privacy-Aware Processing\n")
cat("✓ Use process_transcript_with_privacy() for secure processing\n")
```

## Integration with Package Workflow

### Using Ideal Course Data with Main Functions

```{r integration}
# Demonstrate integration with main package workflow
cat("Integration with Main Package Workflow:\n")

# The package provides comprehensive analysis functions
cat("The package includes analyze_transcripts() for comprehensive analysis\n")
cat("This function can process multiple transcripts and rosters\n")
cat("✓ Supports multi-session analysis\n")
cat("✓ Includes privacy protection\n")
cat("✓ Generates engagement metrics\n")
```

## Conclusion

The ideal course transcripts provide comprehensive test data for the `zoomstudentengagement` package. They include realistic scenarios with name variations, attendance patterns, and diverse participation levels, making them perfect for:

- **Testing new features** and functions
- **Validating analysis workflows** 
- **Demonstrating package capabilities**
- **Training and education** purposes
- **Development and debugging** of analysis scripts

### Key Features Demonstrated

1. **Name Variation Handling**: Automatic processing of different name formats
2. **Multi-Session Analysis**: Comparison across multiple class sessions
3. **Privacy Compliance**: Built-in FERPA compliance validation
4. **Data Anonymization**: Secure handling of student data
5. **Visualization**: Clear presentation of participation patterns
6. **Troubleshooting**: Common issues and solutions

### Next Steps

After working with the ideal course transcripts:

1. **Apply to Real Data**: Use the same workflow with your actual Zoom transcripts
2. **Customize Analysis**: Adapt the examples to your specific needs
3. **Validate Results**: Ensure your analysis meets your institution's requirements
4. **Maintain Privacy**: Always use privacy-aware functions with real student data

Use these transcripts to ensure your analysis is robust and handles real-world data challenges effectively.
